---
title: "Homework 2"
author: Jingyi Zhang
output: github_document
---

```{r setup}
library(tidyverse)
library(readxl)
```

## Problem 1 

Read the Mr. Trashwheel dataset.

```{r}
trashwheel_df = 
  read_xlsx(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "Mr. Trash Wheel",
    range = cell_cols("A:N")) %>%
  janitor::clean_names() %>%
  drop_na(dumpster) %>%
  mutate(
    sports_balls = round(sports_balls),
    sports_balls = as.integer(sports_balls)
  )
```

Read participation data. For 2018 and 2017.

```{r}
precip_2018 = 
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2018 Precipitation",
    skip = 1
  ) %>%
  janitor::clean_names() %>%
  drop_na(month) %>%
  mutate(year = 2018) %>%
  relocate(year)

precip_2017 = 
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2017 Precipitation",
    skip = 1
  ) %>%
  janitor::clean_names() %>%
  drop_na(month) %>%
  mutate(year = 2017) %>%
  relocate(year)
```

Now combine annual precipitation.

```{r}
month_df = 
  tibble(
    month = 1:12,
    month_name = month.name
  )

precip_df = 
  bind_rows(precip_2018, precip_2017)

left_join(precip_df, month_df, by = "month")
```

This dataset contains information from the Mr. Trashwheel trash collector in Baltimore, Maryland. As trash enters the inner harbor, the trashwheel collects that trash, and stores it in a dumpster. The dataset contains information on year, month, and trash collected, include some specific kinds of trash. There are a total of `r nrow(trashwheel_df)` rows in our final dataset. Additional data sheets include month precipitation data. In this dataset:

* The median number of sports balls found in a dumpster in 2017 was `r trashwheel_df %>% filter(year == 2017) %>% pull(sports_balls) %>% median()`.
* The total precipitation in 2018 was `r precip_df %>% filter(year == 2018) %>% pull(total) %>% sum()` inches.

## Problem 2

Read and clean the NYC Transit Subway Entrance and Exit dataset.

```{r}
transit_df =
  read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>%
  janitor::clean_names() %>%
  select(line:entry, vending, ada) %>%
  mutate(entry = recode(entry, "YES" = "TRUE", "NO" = "FALSE", .default = NULL))
```

This dataset contains information on NYC transit. Variables included in this dataset are line, station name, station latitude/longitude, routes served, entry, entrance type, vending, and ADA compliance. This dataset has `r nrow(transit_df)` rows and `r ncol(transit_df)` columns. So far, all column names are converted to lowercase and snakecase form. Then, the uninterested columns are removed from the dataset. The entry variable is converted from character to a logical variable with values of "TRUE" or "FALSE". This dataset at this stage is not tidy yet. The routes served columns (route1-route11) should be combined into one single column.  



How many distinct stations are there?

```{r distinct_stations}
select(transit_df, line, station_name) %>%
  distinct(line, station_name) %>%
  count()
```

How many stations are ADA compliant?

```{r ada_compliant}
filter(transit_df, ada == "TRUE") %>%
  select(ada, line, station_name) %>% 
  distinct(ada, line, station_name) %>% 
  count()
```

What proportion of station entrances/exits without vending allow entrance?

```{r}
filter(transit_df, entry == "TRUE", vending == "NO") %>%
  select(line, station_name, entry, vending) %>%
  distinct(line, station_name, entry, vending) %>%
  count()
43/465
```

Reformat data that makes route number and route name distinct variables.

```{r}
transit_tidy_df = 
  mutate_at(transit_df, vars(route8:route11), as.character) %>%
  pivot_longer(
    route1:route11,
    names_to = "route_name",
    names_prefix = "route",
    values_to = "route_number"
  ) %>%
  drop_na(route_number)
```

How many distinct stations serve the A train?

```{r}
filter(transit_tidy_df, route_number == "A") %>%
  select(route_number, line, station_name) %>%
  distinct(route_number, line, station_name) %>%
  count()
```

Of the stations that serve the A train, how many are ADA compliant?

```{r}
filter(transit_tidy_df, route_number == "A", ada == "TRUE") %>%
  select(route_number, ada, line, station_name) %>%
  distinct(route_number, ada, line, station_name) %>%
  count()
```

## Problem 3

Clean the data in pols-month.

```{r}
pols_month = read_csv("./data/fivethirtyeight_datasets/pols-month.csv") %>%
  separate(mon, into = c("year", "month", "day"), convert = TRUE)
```

First step: Replace month number by month names.

```{r}
month_df =
  tibble(
    month = 1:12,
    month_name = month.name
  )

pols_month =
  left_join(pols_month, month_df, by = "month")

pols_month =
  mutate(pols_month, president = case_when(
    prez_gop == "1" ~ "gop",
    prez_dem == "1" ~ "dem"
  ))
select(pols_month, -day, -prez_gop, -prez_dem, -month) %>%
  relocate(month_name, .after = year)
```

Second step: 

```{r}

```











